A Component-Based Sensor Network for \\ Environmental Monitoring

### 2009-03-14 docs/icccn2009/paper.tex
Modified the Conclusion by removing the first paragraph and adding another that sums up related work.  Supporting bibliography was added.
git-svn-id: svn://localhost/netbeams/trunk@393 559edde2-a4d4-44e0-96fc-4035e40d7396

The California coastal region is populated by a vast number of
remotely deployed sensors that monitor environmental conditions
ranging from weather to water quality to ocean surface currents.
These sensors are operated by data providers, including resource
management groups, research scientists, municipalities, state and
federal agencies, who could use near real-time data streams to inform
decisions that impact public health, public safety, and the protection
of resources. Success of these efforts, whether small scale or a large
integrated network, depends on the first step in the data management
process: reliably obtaining data in near real-time from remotely
located sensors.  This step can be the weak link in the data
management pathway when telemetry options are unreliable.  The
availability of a plug-and-play computing platform designed to
communicate with a variety of sensor types would stabilize data
streams with poor telemetry and allowing expansion of sensor coverage
into regions not presently served.

In this paper, we introduce a practical application of a sensor network
in the domain of environmental monitoring. We have built an
environmental sensor network in the San Francisco Bay using highly
specialized, off-the-shelf water quality sensors. The resulting
infrastructure enables near-realtime access to sensor data collected
from remote locations. Our end-to-end solution, furthermore, allows
configuration and management of remotely deployed sensors. In this sense, the
key to our infrastructure is the DSP, a micro-kernel plug-and-play
architecture that can flexibly be configured to specific needs through
special purpose components. Furthermore, DSP is a contribution to the OSGi
project, providing an optional way of unified loosely-coupled bundle
interaction and distributed communication. The NetBEAMS infrastructure, as
described in the paper, is available as an Open-Source project and can be 
viewed in operation at \texttt{www.netbeams.org}.

>>>>

%The California coastal region is populated by a vast number of
%remotely deployed sensors that monitor environmental conditions
%ranging from weather to water quality to ocean surface currents.
%These sensors are operated by data providers, including resource
%management groups, research scientists, municipalities, state and
%federal agencies, who could use near real-time data streams to inform
%decisions that impact public health, public safety, and the protection
%of resources. Success of these efforts, whether small scale or a large
%integrated network, depends on the first step in the data management
%process: reliably obtaining data in near real-time from remotely
%located sensors.  This step can be the weak link in the data
%management pathway when telemetry options are unreliable.  The
%availability of a plug-and-play computing platform designed to
%communicate with a variety of sensor types would stabilize data
%streams with poor telemetry and allowing expansion of sensor coverage
%into regions not presently served.

In this paper, we introduce a practical application of a sensor network
in the domain of environmental monitoring. We have built an
environmental sensor network in the San Francisco Bay using highly
specialized, off-the-shelf water quality sensors. The resulting
infrastructure enables near-realtime access to sensor data collected
from remote locations. Our end-to-end solution, furthermore, allows
configuration and management of remotely deployed sensors. In this sense, the
key to our infrastructure is the DSP, a micro-kernel plug-and-play
architecture that can flexibly be configured to specific needs through
special purpose components.

Furthermore, our work improves upon existing technologies in earlier sensor network implementations.
Many sensor networks use PIC microcontrollers \cite{Pic01} and specialized operating systems
such as TinyOS \cite{tinyos01}. However PICs are usually too buggy,
requiring non-standard operating systems \cite{Martinez04}. Likewise, TinyOS exhibits limited 
constraints on storage, restricting any complex processing of sensor data
that can be performed \cite{Martinez04}. OSGi has been used in other sensor network
project \cite{Mueller07, Baude07}. Although some OSGi sensor networks also include remote 
communication capabilities \cite{Rellermeyer}, our DSP is designed to support asynchronous 
communication in loosely coupled networks. It is a contribution to the OSGi
project, providing an optional way of unified loosely-coupled bundle
interaction and distributed communication. The NetBEAMS infrastructure, as
described in the paper, is available as an Open-Source project and can be 
viewed in operation at \texttt{www.netbeams.org}.
